---
title: 'CDI Model Systematic Review: Creating tables and figures'
output:
  html_document:
    df_print: paged
  word_document: default
---

Obligatory project downloads and utilities.

```{r, settings}
library(tidyverse)
library(ggthemes)
library(readxl)
library(stringr)
library(kableExtra)
library(flextable)
library(formattable)
library(ggrepel)

# for reproducibility, variable paths are copied below
file_loc <- "~/Desktop/Research/Deshpande Group (C difficile)/CDI Prediction Model Meta Analysis/Systematic Review of Methodology : Reporting/Figures/Analysis"
data_path <- "../figures/2023-01-03_Figure1_tidyData.xlsx"


# set working directory to file source for relative paths
setwd(file_loc)
```

First, read in data from tidied data analysis file.

```{r, data_loading}
# read in tidied data as tibble object
(fig2_data <- as_tibble(read_excel(data_path, 
                                   "predictors")))  # 'predictors' sheet

(fig3_data <- as_tibble(read_excel(data_path, 
                                   "bystudy",       # 'bystudy' sheet
                                   skip = 1)) %>%   # skip header line
   select(auth, n_preds, n_study_train, n_study_total, n_event_total) %>%
   mutate(obs_per_coef = n_study_total / n_preds,   # calc obs / coeff
          events_per_coef = n_event_total / n_preds,# calc evnts / coeff
          validated = !is.na(n_study_train))
  )

fig3_header <- as_tibble(names(read_excel(data_path, 
                                           "bystudy"))) # extract header
```

Now, we create figures 2 and 3 from the manuscript. Figure 1 was generated using a [Shiny App](https://www.eshackathon.org/software/PRISMA2020.html).

- Figure 2 shows the count of predictors in 10 models of C. difficile infection.
- Figure 3 shows the number of events (CDI infection) using a log axis vs. number of model coefficients.

```{r, theme_settings}
# plotting minimal graphics
theme_set(theme_tufte(ticks = F, base_size = 14))
plot_color <- "grey"
```

```{r, fig2}
# figure 2 - count of predictors
ggplot(fig2_data, 
       aes(x = reorder(description, freq), # order by most to least common
           y = freq)) +
  geom_bar(stat="identity", width=0.25, fill = plot_color) +
  coord_flip() +
  scale_y_continuous(breaks=seq(1, 10, 1)) +
  geom_hline(yintercept=seq(1, 10, 1), col="white", lwd=1) +
  geom_hline(yintercept=0, col=plot_color) +
  labs(x = NULL, 
       y = "Number of times included") +
  annotate("text", 
           x = 10.5, 
           y = 10, 
           adj=1,  
           family="serif",
  label = c("Number of occurances of a predictor\nin 10 models of C difficile infection"))
ggsave("figs/fig2.tiff")
```


```{r, fig3}
# figure 3 - number of events (CDI infection) using a log axis vs. number of model coefficients
ggplot(fig3_data, aes(x = n_preds, 
                      y = n_event_total, 
                      label = auth, 
                      shape = validated)
       ) +
  geom_point() +
  labs(y = fig3_header$value[6],        # events (CDI)
       x = fig3_header$value[2]) +      # number of model coefficients
  geom_text_repel(size = 3, 
                  family = "Cambria", 
                  color = 'black', 
                  nudge_x = 0.5) +
  scale_y_continuous(trans='log10') +   # log scale of events (CDI)
  annotation_logticks(sides = "l") +
  theme(legend.position="none") +       # legend explained in figure title
  scale_shape_manual(values=c(17, 19)) +
  annotate("text", 
           x = 20, 
           y = 50, 
           adj=1,  
           family="serif",
  label = c("Triangles denote studies\nwithout a validation set"))
ggsave("figs/fig3.tiff")
```

Tables 1 through 4 are produced next.

```{r, tbl1}
# read in tidied data
(tbl1_data <- as_tibble(read_excel(data_path, 
                                   "bystudy", 
                                   skip = 1))
 )  # NB: looking at predictors sheet

tbl1_data_tidy <- tbl1_data %>%
  relocate(cand_preds, .after = 1) %>%
  mutate(.after = 7, 
         epv = round(n_event_total / cand_preds, 1)) %>% # calc epv
  mutate(.after = 5, 
         spv = round(n_study_total / cand_preds, 1)) %>% # calc spv
  select(1:9) # trim off other columns

tbl1_header <- read_excel(data_path, 
                          "bystudy") %>% 
  names() %>%
  as_tibble() %>%
  add_row(.after = 1, value = "Candidate") %>%
  add_row(.after = 4, value = "per candidate predictor") %>%
  add_row(.after = 7, value = "per candidate predictor") %>%
  slice(1:9)
tbl1_header[c(4,5,7,8),] <- rep.int(c("Derivation set", "Total set"), 2)
tbl1_header[1,] <- " "


# correct for studies with multiple models
tbl1_data_tidy[c(7,8),1] <- as_tibble(c("MGH", "UM"))
tbl1_data_tidy[c(9,10),1] <- as_tibble(c("Index model (2019)", "Validation model (2021)"))
tbl1_data_tidy[c(11,12),1] <- as_tibble(c("Model 1", "Model 2")) 

# change headers to account for additional header rows
tbl1_header[3,] <- "Final"
tbl1_header[6,] <- "per candidate variable"
tbl1_header[9,] <- "per candidate variable"

options(knitr.kable.NA = '')
as_tibble(tbl1_data_tidy) %>%
  kable(escape = F, 
        align = "lcccccccc", 
        format.args = list(big.mark = ","), 
        col.names = tbl1_header$value) %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  pack_rows(paste0("Oh et al. (2018)", footnote_marker_alphabet(1)), 7, 8, escape = F) %>%
  pack_rows(paste0("Tilton et al.", footnote_marker_alphabet(2)), 9, 10, escape = F) %>%
  pack_rows("Voicu et al. (2021)", 11, 12, escape = F) %>%
  kableExtra::footnote(alphabet_title = c("Bolded studies included multiple models"), 
           alphabet = c("Created institution specific models.", "Validated prior model on external population.")) %>%
  add_header_above(c(" " = 1, "Coefficients" = 2, "Samples" = 3, "Events (CDI)" = 3)) %>%
  save_kable(file = "tbls/tbl1.pdf", density = 300)
```

```{r, tbl2}
as_tibble(read_excel(data_path, 
                     "t3", 
                     skip = 1)) %>%
  select(1:4) %>%
  kable(align = "lccc", col.names = names(as_tibble(read_excel(data_path, "t3")[1,1:4]))) %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  pack_rows(paste0("Oh et al. (2018)", footnote_marker_alphabet(1)), 7, 8, escape = F) %>%
  pack_rows(paste0("Tilton et al.", footnote_marker_alphabet(2)), 9, 10, escape = F) %>%
  pack_rows("Voicu et al. (2021)", 11, 12, escape = F) %>%
  column_spec(1, width_min = "13em") %>%
  column_spec(4, width_min = "13em") %>%
  save_kable(file = "tbls/tbl2.pdf", density = 800)
```

```{r, tbl3}
# Table 3 – Evaluating performance of risk prediction models
## load in data for t3
(tbl3_data <- as_tibble(read_excel(data_path, 
                                 "t2", 
                                 skip = 1)) %>%
   select(-c("cal", "Live models")) %>%          # drop columns with no data
   mutate(across(-1, ~ replace_na(.x, FALSE)))
) 

# group by column occurrence (exclude author)
tbl3_data %>% 
  summarize(across(-auth, sum)) %>%
  pivot_longer(cols = 1:8, names_to = "Parameter", values_to = "Sum") %>%
  add_column(.before = "Sum", desc = names(read_excel(data_path, 
                                                      "t2")[1,2:9])) %>%
  mutate(.after = "Sum",
         perc = paste0(round((Sum / nrow(tbl3_data)), 2)*100, "%")) %>%
  mutate(disp = paste0(Sum, " (", perc,")")) %>%
  select(desc, disp) %>%
  kable(align = "lc", col.names = c("Parameter", 
                                    "Number of models (% of total)")) %>%
  kable_classic(full_width = F, 
                html_font = "Cambria") %>%
  pack_rows("Validation", 1, 4) %>%
  pack_rows("Performance metrics", 5,6) %>%
  pack_rows("Calibration", 7,8) %>%
  save_kable(file = "tbls/tbl3.pdf", density = 800)
```

```{r, tbl4}
tbl4_data <- as_tibble(read_excel(data_path, 
                                   "t1", 
                                   skip = 1))

tbl4_header <- as_tibble(read_excel(data_path, 
                                    "t1")[1,])
names(tbl4_header)[1] <- " "

tbl4_data[c(7,8),1] <- as_tibble(c("MGH", "UM"))
tbl4_data[c(9,10),1] <- as_tibble(c("Index model (2019)", "Validation model (2021)"))
tbl4_data[c(11,12),1] <- as_tibble(c("Model 1", "Model 2"))

as_tibble(tbl4_data) %>%
  select(1,8:9) %>%
  kable(escape = F, align = "lccc", format.args = list(big.mark = ","), 
        col.names = c("", "Derivation", "Validation")) %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  pack_rows(paste0("Oh et al. (2018)", footnote_marker_alphabet(1)), 7, 8, escape = F) %>%
  pack_rows(paste0("Tilton et al.", footnote_marker_alphabet(2)), 9, 10, escape = F) %>%
  pack_rows("Voicu et al. (2021)", 11, 12, escape = F) %>%
  kableExtra::footnote(general = c("Performance was reported as by area under the ROC for all studies.", "Values rounded to two decimal places.", "NR = Not reported"),
          alphabet_title = c("Bolded studies included multiple models"), 
          alphabet = c("Created institution specific models.", "Validated prior model on external population.")) %>%
  add_header_above(c(" " = 1, "Performance" = 2)) %>%
  column_spec(1, width_min = "13em") %>%
  save_kable(file = paste0("tbls/tbl4.pdf"), density = 300)
```